---
title: "homework-week-6"
author: "Maria C. Codlin"
date: "October 17, 2017"
output: html_document
---
[1] Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines.
Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().

When conducting a two-sample test, it should be *p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”*, the same as in the use of x and y in the function t.test().

The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.

The function should contain a check for the rules of thumb we have talked about (n∗p>5n∗p>5 and n∗(1−p)>5n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings.

If this is violated, the function should still complete but it should also print an appropriate warning message.

The function should return a list containing: 
    the members Z (the test statistic), 
    P (the appropriate p value), and 
    CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). 
    For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.

```{r}
Z.prop.test<-function(p1, n1, p2=NULL, n2=NULL, p0, alternative="two.sided", conf.level=0.95){
  
  
#tests
  if(is.null(p2)||is.null(n2)){
    z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
  }
if ((n1∗p0>5) & n1∗(1−p0)>5) {
  #runtest
} else {#runtest 
  paste("Warning: failed one or more validity tests, data may not be normally distributed")
  }
```



```{r}
if (n1*p1>5) & (n2*p2>5) & n1(1-p1)>5 & n2(1-p2)>5{
  #runtest
} else {#runtest 
  paste("Warning: failed one or more validity tests, data may not be normally distributed")
  }

  
#one sample
  z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
  
#two sample
  pstar<-(p1 * n1 + p2 * n2) / (n1 + n2)
  z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2))
  
  if (alternative="greater"){
    if (p2>p1) {}
  }
  ]
  if (alternative="greater"){
    if (p2<p1){}
  }
  ]
}
```

