---
title: "homework-week-6"
author: "Maria C. Codlin"
date: "October 17, 2017"
output: html_document
---
[1] Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines.
Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().

When conducting a two-sample test, it should be *p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”*, the same as in the use of x and y in the function t.test().

The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.

The function should contain a check for the rules of thumb we have talked about (n∗p>5n∗p>5 and n∗(1−p)>5n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings.

If this is violated, the function should still complete but it should also print an appropriate warning message.

The function should return a list containing: 
    the members Z (the test statistic), 
    P (the appropriate p value), and 
    CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). 
    For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.


```{r function part 1, one sample prop test}
Z.prop.test<-function(p1, n1, p2=NULL, n2=NULL, p0, alternative="two.sided", conf.level=0.95) {
  
    if(is.null(p2)||is.null(n2)) {
      
  #one sample test      
if ((n1∗p1<=5) | n1∗(1−p1)<=5) warning("Warning: failed one or more validity tests, data may not be normally distributed") #in the module, this is given here refering to pi, not p-hat, but Davies (2016) writes this with p-hat.      
    if (alternative == "two.sided") {
      z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
      p <- 1 - pnorm(z, lower.tail = TRUE) + pnorm(z, lower.tail = FALSE)
    lower <- p1 - qnorm((1 - (1 - conf.level)/2)) * sqrt(p1 * (1 - p1)/n1)
    upper <- p1 + qnorm((1 - (1 - conf.level)/2)) * sqrt(p1 * (1 - p1)/n1)
    ci <- c(lower, upper)
    }
  
    if (alternative =="greater") {
      if (p1<p0) stop ("p0>p1, use alternative='less'")
    z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
    p <- pnorm(z, lower.tail=FALSE)
    lower <- p1 - qnorm((1 - (1 - conf.level)/2)) * sqrt(p1 * (1 - p1)/n1)
    upper <- p1 + qnorm((1 - (1 - conf.level)/2)) * sqrt(p1 * (1 - p1)/n1)
    ci <- c(lower, upper)
    }
    
    if (alternative =="less") {
      if (p1>p0) stop ("p1>p0, use alternative='greater'")
    z <- (p1-p0)/sqrt(p0 * (1 - p0)/n1)
    p <- pnorm(z, lower.tail=TRUE)
    lower <- p1 - qnorm((1 - (1 - conf.level)/2)) * sqrt(p1 * (1 - p1)/n1)
    upper <- p1 + qnorm((1 - (1 - conf.level)/2)) * sqrt(p1 * (1 - p1)/n1)
    ci <- c(lower, upper)
    }
    }
  else{
  #two sample test
    
    if ((n1∗p1<=5) | n1∗(1−p1)<=5 | (n2∗p2<=5) | n2∗(1−p2)<=5) warning ("Warning: failed one or more validity tests, data may not be normally distributed") #in the module, this is given here refering to pi, not p-hat, but Davies (2016) writes this with p-hat.
    
    if (alternative == "two.sided") {
  pstar<-(p1 * n1 + p2 * n2) / (n1 + n2)
  z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2))
  p <- 1 - pnorm(z, lower.tail = TRUE) + pnorm(z, lower.tail = FALSE)
  lower <- (p2 - p1) - qnorm((1 - (1 - conf.level)/2)) * sqrt((p1*(1-p1))/n1 + (p2*(1 - p2))/(n2))
  upper <- (p2 - p1) + qnorm((1 - (1 - conf.level)/2)) * sqrt((p1*(1-p1))/n1 + (p2*(1 - p2))/(n2))
  ci <- c(lower, upper)
    }
  
    if (alternative =="greater") {
      if (p1>p2) stop ("p2>p1, use alternative='less'")
   
  pstar<-(p1 * n1 + p2 * n2) / (n1 + n2)
  z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2))
  p <- 1 - pnorm(z, lower.tail = FALSE)
  lower <- (p2 - p1) - qnorm((1 - (1 - conf.level)/2)) * sqrt((p1*(1-p1))/n1 + (p2*(1 - p2))/(n2))
  upper <- (p2 - p1) + qnorm((1 - (1 - conf.level)/2)) * sqrt((p1*(1-p1))/n1 + (p2*(1 - p2))/(n2))
  ci <- c(lower, upper)
      
    }
    
  if (alternative =="less") {
    if (p1<p2) stop ("p1>p2, use alternative='greater'")
      pstar<-(p1 * n1 + p2 * n2) / (n1 + n2)
      z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2))
      p <- 1 - pnorm(z, lower.tail = TRUE)
      lower <- pstar - qnorm((1 - (1 - conf.level)/2)) * sqrt(pstar * (1 - pstar)/(n1+n2))
      upper <- pstar + qnorm((1 - (1 - conf.level)/2)) * sqrt((p1*(1-p1))/n1 + (p2*(1 - p2))/(n2))
      ci <- c(lower, upper)
    }}
  
    l<-list(
    "Z" = z,
    "p" = p,
    "Confidence interval" = ci)
  return (l) 
    }


```

```{r}
Z.prop.test(p1=0.56, p2=0.7, n1=25, n2=30, alternative="two.sided")
Z.prop.test(p1=0.56, p2=0.7, n1=25, n2=30, alternative="greater")
Z.prop.test(p1=0.56, p2=1, n1=25, n2=30, alternative="greater")
```
```{r}
Z.prop.test(p1=0.6, p0=0.8, n1=30)
Z.prop.test(p1=0.9, p0=0.8, n1=10, alternative = "greater")
```

```
**TO DO**

- add in additional warnings

```{r}
Z.prop.test<-function(p1, n1, p2=NULL, n2=NULL, P0, alternative="two.sided", conf.level=0.95) {
  if(is.null(p2)||is.null(n2)) {
#one sample test
    if (alternative == "two.sided") {
      z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
      p <- 1 - pnorm(z, lower.tail = TRUE) + pnorm(z, lower.tail = FALSE)
          lower <- p1 - qnorm(0.975) * sqrt(p1 * (1 - p1)/n1)
    upper <- P1 + qnorm(0.975) * sqrt(P1 * (1 - P1)/n1)
    ci <- c(lower, upper)
      
    if (P0<p1)#greater
    z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
    p <- pnorm(z, upper.tail=TRUE)
    lower <- p1 - qnorm(0.975) * sqrt(p1 * (1 - p1)/n1)
    upper <- P1 + qnorm(0.975) * sqrt(P1 * (1 - P1)/n1)
    ci <- c(lower, upper)
    
     if (P0>p1)#lesser
    z <- (p1-p0)/sqrt(p0 * (1 - p0)/n1)
    p <- pnorm(z, lower.tail=TRUE)
    lower <- p1 - qnorm(0.975) * sqrt(p1 * (1 - p1)/n1)
    upper <- P1 + qnorm(0.975) * sqrt(P1 * (1 - P1)/n1)
    ci <- c(lower, upper)
    
    l<-list(
    "Z" = z,
    "p" = p,
    "Confidence interval" = ci)
    }
    
    
  }
  
  if ((n1∗p0<=5) & n1∗(1−p0)<=5) {
    paste("Warning: failed one or more validity tests, data may not be normally distributed")}
 return(l) 
}

```

```{r}
Z.prop.test(P1=0.6, P0=0.8, N1=30)
```



```{r}
if (n1*p1>5) & (n2*p2>5) & n1(1-p1)>5 & n2(1-p2)>5{
  #runtest
} else {#runtest 
  paste("Warning: failed one or more validity tests, data may not be normally distributed")
  }

  list(
    "Z" = z
    "p" = 
    "Confidence interval" = 
  )
#one sample
  z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1)
  
#two sample
  pstar<-(p1 * n1 + p2 * n2) / (n1 + n2)
  z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2))
  
  if (alternative="greater"){
    if (p2>p1) {}
  }
  ]
  if (alternative="greater"){
    if (p2<p1){}
  }
  ]
}
```

test